---
Assignment: Final Project
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

## Load Datasets
```{r}
salary_data <- read_csv("data/salary_data.csv")
pollution_data <- read_csv("data/cities_air_quality_water_pollution.18-10-2021.csv.xls")
head(salary_data)
head(pollution_data)
```

## Question 1: How does the level of pollution vary in regions with different levels of income? Can income predict pollution for a given area?

## Pick out Columns We Need
```{r}
salary_data <- salary_data %>% 
  select(country_name, median_salary, average_salary) %>% 
  rename(country = country_name)

pollution_data <- pollution_data %>% 
  select(Country, AirQuality, WaterPollution) %>% 
  rename(country = Country) %>% 
  rename(air_quality = AirQuality) %>% 
  rename(water_pollution = WaterPollution)

head(salary_data)
head(pollution_data)
```

## Adjust Pollution Data to 1 Per Country
```{r}
pollution_data <- pollution_data %>% 
  drop_na(air_quality) %>% 
  drop_na(water_pollution)

pollution_data <- pollution_data %>% 
  arrange(country) %>% 
  group_by(country) %>% 
  summarize(air_quality = mean(air_quality, na.rm = TRUE),
            water_pollution = mean(water_pollution, na.rm = TRUE))
head(pollution_data)
```

## Merge Datasets
```{r}
combined_data <- left_join(salary_data, pollution_data)

num_nas <- combined_data %>% 
  select(air_quality) %>% 
  is.na() %>% 
  sum() %>% 
  print()

combined_data <- left_join(salary_data, pollution_data) %>% 
  drop_na(air_quality)
combined_data
```

## Graph Relationship Between Salaries and Air Quality
```{r}
combined_data %>% 
  select(median_salary, air_quality) %>% 
  ggplot(aes(x = median_salary, y = air_quality, color = median_salary)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen")  +
  labs(x = "Median Salary",
       y = "Air Quality",
       title = "Relationship Between Median Salary and Air Quality")
```

```{r}
combined_data %>% 
  select(average_salary, air_quality) %>% 
  ggplot(aes(x = average_salary, y = air_quality, color = average_salary)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen") +
  labs(x = "Average Salary",
       y = "Air Quality",
       title = "Relationship Between Average Salary and Air Quality")
```

## Graph Relationship Between Salaries and Water Pollution
```{r}
combined_data %>% 
  select(median_salary, water_pollution) %>% 
  ggplot(aes(x = median_salary, y = water_pollution, color = median_salary)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Median Salary",
       y = "Water Pollution",
       title = "Relationship Between Median Salary and Level of Water Pollution")
```

```{r}
combined_data %>% 
  select(average_salary, water_pollution) %>% 
  ggplot(aes(x = average_salary, y = water_pollution, color = average_salary)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Average Salary",
       y = "Water Pollution",
       title = "Relationship Between Average Salary and Level of Water Pollution")
```

## Question 2: Do air pollution and water pollution always overlap? Is air pollution a good predictor of water pollution and vice versa?

## Load Dataset
```{r}
pollution_data2 <- read_csv("data/cities_air_quality_water_pollution.18-10-2021.csv.xls")
head(pollution_data2)
```

## Select Columns
```{r}
pollution_data2 <- pollution_data2 %>% 
  select(AirQuality, WaterPollution) %>% 
  rename(air_quality = AirQuality) %>% 
  rename(water_pollution = WaterPollution) 
head(pollution_data2)
```

## Filter Anomalies
```{r}
pollution_data2 <- pollution_data2 %>% 
  drop_na(air_quality) %>% 
  drop_na(water_pollution)

Q1 <- quantile(pollution_data2$air_quality, 0.25)
Q3 <- quantile(pollution_data2$air_quality, 0.75)
IQR_value1 <- Q3 - Q1

Q1 <- quantile(pollution_data2$water_pollution, 0.25)
Q3 <- quantile(pollution_data2$water_pollution, 0.75)
IQR_value2 <- Q3 - Q1
```

## Graph Their Relationship
```{r}
pollution_data2 %>% 
  filter(air_quality > (Q1 - 1.5 * IQR_value1) & air_quality < (Q3 + 1.5 * IQR_value1),
         water_pollution > (Q1 - 1.5 * IQR_value2) & water_pollution < (Q3 + 1.5 * IQR_value2)) %>% 
  ggplot(aes(x = air_quality, y = water_pollution)) +
  geom_smooth(method = "lm", se = FALSE, color = "darkred") +
  geom_point() +
  labs(x = "Air Quality", 
       y = "Water Pollution", 
       title = "Relationship Between Air Quality and Water Pollution")
```